The Spatial and Temporal Domains of Modern Ecology

### 2017-06-12 paper/main.tex
Everything through final discussion section revised

\usepackage[sort]{scicite}
\usepackage{times}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{color}

@@

%\begin{sciabstract}
\section*{Abstract}
%Ecologists have long known that the observational scales play a critical role in shaping our understanding of ecosystems, and that observations spanning multiple scales of space and time are needed to effectively characterize ecological phenomena. 
\textbf{In order to properly understand ecological phenomena, it is necessary to quantify their behavior over the range of spatial and temporal scales at which they manifest. Ecology has been concerned with this need since the early 1990s, and the ability to collect multi-scaled ecological observations has grown rapidly since then.  Characterizing the spatial and temporal domains of modern ecological observations can therefore provide important insight into the field's progress in understanding towards a more comprehensive understanding of ecosystem behaviour. To characterize these domains, we conducted a meta-analysis of recent (2004-2014) ecological studies, in which we quantified four primary dimensions of their reported observations: plot resolution, sampling interval, effective duration (time between start and end of temporal replicates), and effective extent (area enclosed by spatial replicates). We also estimated the \emph{actual} extent and duration, which respectively represent the summed area and time covered by spatial and temporal replicates. \textcolor{blue}{Replace this text with more specific summary of observed scales:} \textcolor{red}{Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}}}. % Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations). Resolution, extent, and duration were not clearly reported in most reviewed studies. This indicates inattentiveness to scale, which, together with the narrow spatio-temporal domains of observations, suggests a limited ability to generalize the knowledge gained in many ecological studies, which is critical for developing an ability to predict ecological responses in an era of increasing global change \cite{levin_problem_1992}. An emerging wave of observing and analytical technologies (e.g. nano-satellites, unmanned systems, open image archives, and high-power processing platforms) now makes it dramatically easier and cheaper to collect high-quality, ecologically meaningful measurements at scales inaccessible to field scientists. Combined with a greater attention to scale in research design, these advances can facilitate the spread of ecological research to sparsely observed scales, while improving the transferability of ecological knowledge.}

@@

\linenumbers
\vspace{10pt}
The scales at which ecosystems are observed plays a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.  Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}. This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad range of scales \cite{schneider_rise_2001, chave_problem_2013}.

Given this awareness of the centrality of scale to understanding ecology, and the growing ability to study ecological phenomena across a broad range of spatial and temporal scales, it is important to assess the scales at which current ecological research is conducted. To gain insight into this question, we quantified the spatial and temporal domains of empirical observations that were reported in a representative sample of studies published between 2004-2014 in the top 30 ecological journals (by impact factor). Empirical observations provide the necessary means for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, thus it stands to reason that the temporal and spatial distributions of ecological observations may be indicative of modern ecology's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.

We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, interval and duration (Table 1).  Here resolution is the area of an individual spatial replicate, or the two-dimensional space in which all measurable features of a natural object(s) were recorded (as opposed to sub-sampled), while extent is the area enclosed by the outer-most spatial replicates, or, if the system or habitat being sampled was distinct from its surrounding matrix (e.g. forest patches in grassland habitats), the summed area of sampled patches (see SI for full definition). Interval refers to the average time elapsed between individual temporal replicates, and duration is the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation (SI). We also calculated two additional metrics, the integrated area of spatial replicates (i.e. resolution multiplied by number of replicates) and the summed observational time of all temporal replicates. We estimated these additional dimensions to evaluate the degree to which the actual scales of ecological observations differ from those they ostensibly represent, and therefore refer to them as the \emph{actual} extent and \emph{actual} duration.

\begin{table}[htp]
\caption{The dimensions of ecological observations estimated in this meta-analysis.}
\begin{center}
\begin{tabular}{ll}
\hline
\textbf{Dimension} & \textbf{Description} \\
\hline
Resolution & Area (m$^2$) of an individual spatial replicate (plot) \\
Extent & Area (ha) encompassed by all spatial replicates \\
Actual extent & Summed area (ha) of all spatial replicates\\
Interval & Time elapsed (days) between successive temporal replicates \\
Duration & Time elapsed (days) between first and last temporal replicates\\
Actual duration & Summed observational time (days) of all temporal replicates\\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}%

We calculated these dimensions from 379 discrete observations reported within a 134 paper subset of 348 randomly selected articles (from 42,918 total). An additional 62 papers that were cited as the source of observations in the selected papers were also reviewed. We confined our analysis to observations made of ``natural'' \cite{tilman_ecological_1989}, or non-experimentally manipulated, systems, given that the inclusion of experiments could have skewed our assessments towards the relatively fine scales at which such studies typically focus.

To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI). We constructed histograms for each dimension from the mean of the perturbed ensembles, and estimated 95\% confidence intervals for each histogram bin (Fig. 1). We constructed kernel density estimates from the full resampled ensemble in order to assess observational distributions within different juxtapositions of the four space-time dimensions (Fig. 2).

In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig. 1A). The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 42\% 1000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig. 1B).

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/hists_bs.pdf}
\vspace{-0.15 cm}
\caption{Histograms of the resolution (A), extent (B), interval (C), and duration (D) of observations collected from the surveyed ecological studies. Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval. }
\label{afoto1}
\end{figure}

@@

Juxtaposing these observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (the 37\% that were unreplicated were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident. This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations. This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/kde45.pdf}
\vspace{-0.15 cm}
\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations). Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.}
\label{afoto1}
\end{figure}

@@

Comparing the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution that have extents ranging between slightly over 1,000 to nearly 1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha.  These three concentrations suggest a tendency for observational extent to increase with resolution, which is a relationship that becomes more pronounced in the (less densely observed) portion of the domain where resolutions $\geq$100 m$^2$.

A similar tendency for duration to increase with interval was also evident amongst temporally replicated observations (Fig. 2D), where the majority of observations were repeated at daily to decadal intervals and spanned  $\geq$ 1 month to $<$100 years.  The orientation of this concentration shows that observations lasting one year to one decade tend to have corresponding intervals, while those lasting one month to one year have daily to monthly intervals. This suggests that most long-duration observations have just a single temporal replicate. The low densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken.

To provide further insight into the spatio-temporal domains of ecological observations, we also evaluated the differences in scale between extent and actual extent and between duration and actual duration. To make these comparisons, we calculated the magnitudes of difference (decades) between each dimension and its actual equivalent, and examined how these varied as a function of the scale of the actual measurement (Fig. 3). The majority (80\%) of the assessed observations had actual extents of $\leq$1 ha, which on average was 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).  Actual extent converged with extent above 100,000 ha, but this applies to just 5\% of observations. The actual duration of 65\% of observations is $<$ 1 day, which is on average 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig. 3B). The two duration measures were effectively the same for the 16\% of observations exceeding one month of actual duration.

@@

\noindent \textbf{\emph{Observational methods}}\\
We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches. Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.3\%, and paleo-reconstruction and other geographic data each for less than 1\%. Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance. Automated sensing methods showed no trend (SI).

\noindent \textbf{\emph{Potential biases and uncertainties in quantifying observational scales}}\\
%This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI). Observations that were non-continuous in time or space (e.g. point-based field measurements) may effectively represent larger scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}. From a spatial perspective, this concern does not apply to contiguous sampling schemes, which were primarily based on remote sensing and cover the largest area, thus the net effect is that this analysis may underestimate the effective areas of smaller observational extents. From a temporal perspective, our analysis likely underestimates the effective duration of many observations, particularly those where instantaneous, repeated measures of slow-changing ecological features were made. Snapshots in time may be sufficient to capture the temporal dynamics of such phenomena (e.g. changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these the total time elapsed between the first and last measurements (the span of observation) may more closely approximate effective duration. However, many studies focused on more dynamic phenomena, and for these long periods of continuous observation may be more important for understanding dynamics than frequent repeats. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to directly measure fire behavior \cite{clements_observing_2007}, except patterns resolving themselves at scales greater than 12-24 hours. For this reason, and to provide a consistent standard for estimating duration, we recorded actual duration.

%Adding to our motivation to record actual observational scales was the fact that 
There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains. The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate, rather than simply record, these values for most observations (specifically, in 63, 60, and 67\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration). The inevitable errors caused by estimation may have biased our overall findings. However, we attempted to quantify this error by assessing inter-rater disagreement and resampling techniques. The resulting confidence intervals (Fig. 1) suggest that it was unlikely that estimation errors unduly influenced our findings.

Another potential source of bias was the rule set we used to estimate observational scales.

Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates. Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find that remote sensing was used for 7.7\% of observations (a 22\% increase), which would increase the mean effective extent by 17.4\% (95\% CI = -1.3-67\%), or 0.07 orders of magnitude, above that observed here (see SI for details of calculation). Further support for such trend can be found by averaging dimensions by publication year (SI), which reveals that observation extent increased by 0.253 orders of magnitude per year between 2004-2014 (R$^2$ = 0.2.5, p$<$0.07). This somewhat clearer trend also suggests that including more recent studies would show observational extent to be somewhat larger, although in this case by a more modest 5.5\% (0.02 orders of magnitude).

@@

\noindent \textbf{\emph{Insights into the scales of modern ecology}}\\
%These results indicate that for most ecological observations are collected with spatially discontinuous replicates that sample very small portions of the area in which they are distributed.

the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.

satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.

Adapt and tone down, retain these points:
Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}. (Scale is not a concern that cuts across the entire discipline of ecology). 
Narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) 
Ecological understanding drawn from many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}, 
a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}.

>>>>

\usepackage[sort]{scicite}
\usepackage{times}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{color}
\usepackage{multirow}

@@

%\begin{sciabstract}
\section*{Abstract}
%Ecologists have long known that the observational scales play a critical role in shaping our understanding of ecosystems, and that observations spanning multiple scales of space and time are needed to effectively characterize ecological phenomena. 
\textbf{In order to properly understand ecological phenomena, it is necessary to quantify their behavior over the range of spatial and temporal scales at which they manifest. Ecology has been concerned with this need for decades, and the ability to collect ecological observations across multiple scales has grown rapidly since then. Characterizing the spatial and temporal domains (the distribution of observations within one or more scale spectra) of modern ecological observations can therefore provide important insight into the field's progress towards a more comprehensive understanding of ecosystem behaviour. To characterize these domains, we conducted a meta-analysis of recent (2004-2014) ecological studies, in which we quantified four primary dimensions of their reported observations: plot resolution, sampling interval, effective duration (time between start and end of temporal replicates), and effective extent (area enclosed by spatial replicates). We also estimated the \emph{actual} extent and duration, which respectively represent the summed area and time covered by spatial and temporal replicates. \textcolor{blue}{Replace this text with more specific summary of observed scales:} \textcolor{red}{Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}}}. % Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations). Resolution, extent, and duration were not clearly reported in most reviewed studies. This indicates inattentiveness to scale, which, together with the narrow spatio-temporal domains of observations, suggests a limited ability to generalize the knowledge gained in many ecological studies, which is critical for developing an ability to predict ecological responses in an era of increasing global change \cite{levin_problem_1992}. An emerging wave of observing and analytical technologies (e.g. nano-satellites, unmanned systems, open image archives, and high-power processing platforms) now makes it dramatically easier and cheaper to collect high-quality, ecologically meaningful measurements at scales inaccessible to field scientists. Combined with a greater attention to scale in research design, these advances can facilitate the spread of ecological research to sparsely observed scales, while improving the transferability of ecological knowledge.}

@@

\linenumbers
\vspace{10pt}
The scales at which ecosystems are observed plays a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.  Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992}. This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad and increasing range of scales \cite{schneider_rise_2001, chave_problem_2013}.

%One of the best known and most comprehensive assessments of scale in ecological research ident

Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) analysis of ecological research scales was published in 1989 \cite{tilman_ecological_1989}, it is both timely and important to assess the scales of contemporary ecological investigation. To address this need, we quantified the spatial and temporal domains of empirical observations that were reported within recently (2004-2014) published ecological studies (here domain means the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}). Empirical observations are critical for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, therefore the spatio-temporal domains of observations provide an important indicator of the field's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.

%This question may be best addressed by evaluating the scales of empirical observations. Empirical observations are critical for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, therefore the spatio-temporal domains of observations provide an important indicator of the field's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.

Our analysis focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1). Resolution is the area of an individual spatial replicate within which a complete measurement (as opposed to a sub-sample) of the feature of interest was made. Extent is the area enclosed by the outer-most spatial replicates, or, if the system or habitat being sampled was distinct from its surrounding matrix (e.g. forest patches in grassland habitats), the summed area of sampled patches (see SI for full definition). Interval refers to the average time elapsed between individual temporal replicates. Duration measures the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation (SI). We also assessed observational scales within two additional dimensions, \emph{actual} extent (the integrated area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates). We evaluated these additional dimensions to gain insight into how much the actual scales of observation (i.e. how much space and time is covered by the measurement) differ from those they are intended to represent.

%For example, consider a hypothetical study reporting estimates of herbaceous plant cover in a forest understory, in which the authors reported that they placed 10 400 m$^2$ plots throughout the forest, and within each plot randomly placed 10 1 m$^2$ quadrats, within which they estimated percent herbaceous plant cover. In this example, we consider 1 m$^2$ to be the resolution.

\begin{table}[htp]
\caption{The scale dimensions of ecological observations assessed in this meta-analysis.}
\begin{center}
\begin{tabular}{llll}
\hline
\multicolumn{2}{c}{\textbf{Component}} & \textbf{Units} & \textbf{Description} \\\hline\hline
\multirow{3}*{Spatial} & Resolution & m$^2$ & Area of an individual spatial replicate (plot) \\%\cmidrule(l){2-3}
   & Extent & ha & Area  encompassed by all spatial replicates \\
   & Actual extent & ha & Summed area of all spatial replicates\\\hline
\multirow{3}*{Temporal} & Interval & days & Time elapsed between successive temporal replicates \\
   & Duration & days & Time elapsed between first and last temporal replicates\\
   & Actual duration & days & Summed observational time of all temporal replicates\\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}%

%We extracted these scale dimensions from 379 observations that were reported in 134 papers that were published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals. 
%

Our analysis was based on a review of 348 papers randomly selected from 42,918 published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals. We extracted scale data from 379 observations of ``natural'' (i.e. non-experimentally manipulated) ecological features that were reported within 134 of the reviewed papers (plus an additional 62 that these cited as the source of observations). We excluded experiments because they tend to be of limited extent, duration, and resolution due to their higher logistical costs \cite{tilman_ecological_1989, kareiva_spatial_1988}, and would therefore likely bias our findings towards finer scales.

%In contrast, cost is less likely to dictate the choice of observational scale in studies of natural systems (considering, for example, that many global-scale, multi-decadal, satellite-derived datasets are free \cite[e.g.][]{tucker_nasas_2004,friedl_modis_2010}), therefore an assessment confined to such studies may provide a less biased picture of ecology's scale domains.

%We calculated these from 379 discrete observations reported within a 134 paper subset of 348
%randomly selected articles (from 42,918 total). An additional 62 papers that were cited as the
%source of observations in the selected papers were also reviewed. We confined our analysis to observations of “natural” (7), or non-experimentally manipulated, phenomena. We made this choice because experiments are costly and therefore tend to be fine-scaled in both space and time (7, 9), which could unduly bias our analysis.
%
%We extracted these scale dimensions from 379 observations reported within 134 papers selected from the top 30 (based on impact factor) ecology-themed journals. The papers were a subset of 348 that we reviewed in total, which we selected randomly from the 42,918 studies published between 2004-2014 . An additional 62 papers that were cited as the source of observations in the selected papers were also reviewed. We confined our analysis to observations of ``natural'' \cite{tilman_ecological_1989}, or non-experimentally manipulated, phenomena. We made this choice because experiments are costly and therefore tend to be fine-scaled in both space and time \cite{tilman_ecological_1989,kareiva_spatial_1988}, which could unduly bias our analysis.

To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI). We constructed histograms for each dimension from the mean of the perturbed ensembles, and estimated 95\% confidence intervals for each histogram bin (Fig. 1). We constructed kernel density estimates from the full resampled ensemble in order to assess observational distributions within different juxtapositions of the four primary (resolution, extent, interval, duration) space-time dimensions (Fig. 2).

\noindent \textbf{\emph{Domains}}\\
In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig. 1A). The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 42\% 1000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig. 1B).

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/hists_bs.png}
\vspace{-0.15 cm}
\caption{Histograms of the resolution (A), extent (B), interval (C), and duration (D) of observations collected from the surveyed ecological studies. Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval. }
\label{afoto1}
\end{figure}

@@

Juxtaposing these observational dimensions provides further insight into the spatio-temporal domains of ecological observations (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (the 37\% that were unreplicated were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident. This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations. This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.

@@

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/kde45.pdf}
\vspace{-0.15 cm}
\caption{Kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations). Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.}
\label{afoto1}
\end{figure}

@@

Comparing the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution that have extents ranging between slightly over 1,000 to nearly 1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha.  These three concentrations suggest a tendency for observational extent to increase with resolution, which is a relationship that becomes more pronounced in the (less densely observed) portion of the domain where resolutions $\geq$100 m$^2$.

A similar tendency for duration to increase with interval was also evident amongst temporally replicated observations (Fig. 2D), with the majority of repeated at daily to decadal intervals and spanning $\geq$ 1 month to 1 decade.  The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting one year to one decade tend to have yearly to decadal intervals. The low densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken.

To provide further insight into observational domains, we also evaluated the degree to which the scales \emph{represented} by observations differed from those that were actually observed. To make this assessment, we first log$_10$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, in order to calculate the magnitude of difference (or decade) between each pair of dimensions for each observation. We then compared how these magnitudes varied in relation to the scale of the actual dimension (Fig. 3). This comparison showed that a majority (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).  Actual extent converged with extent above 100,000 ha, but this applies to just 5\% of observations. The actual duration of 64\% of observations was $<$ 1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig. 3B). The two duration measures were approximately equal for the 16\% of observations exceeding one month of actual duration.

@@

\noindent \textbf{\emph{Observational methods}}\\
We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches. Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.3\%, and paleo-reconstruction and other geographic data each for less than 1\%. Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance. Automated sensing methods showed no trend over time (SI).

\noindent \textbf{\emph{Potential biases and uncertainties in quantifying scales}}\\
There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains. The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate--rather than simply record--these values for most observations (specifically, in 63, 60, and 67\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively). The inevitable estimation errors may have biased our overall findings. However, we attempted to quantify this error by assessing inter-rater disagreement and incorporating this uncertainty into our resampling methodology. The resulting confidence intervals (Fig. 1) suggest that it was unlikely that estimation errors unduly influenced our findings.

Another potential source of bias lies within our scale-estimation protocols, chiefly with respect to our rule for estimating resolution (the smallest areal unit of \emph{complete} measurement). We selected this definition for the sake of consistency, but some papers reported resolution as a larger area in which sub-samples were taken. For these, our estimates are finer than what the studies' authors apparently considered to be plot resolution. Additionally, our domain estimates would presumably be somewhat different if we had included experimentally manipulated observations. For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}.

Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates. Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average (see SI for details of calculation). Further evidence for this trend lies within the extent values themselves, which increased 0.25 orders of magnitude per year between 2004-2014 (R$^2$ = 0.25, p$<$0.07). This somewhat clearer trend also suggests that including more recent studies would show extent to be somewhat larger, although in this case by a more modest 5.5\% (0.02 orders of magnitude).

@@

%This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI). Observations that were non-continuous in time or space (e.g. point-based field measurements) may effectively represent larger scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}. From a spatial perspective, this concern does not apply to contiguous sampling schemes, which were primarily based on remote sensing and cover the largest area, thus the net effect is that this analysis may underestimate the effective areas of smaller observational extents. From a temporal perspective, our analysis likely underestimates the effective duration of many observations, particularly those where instantaneous, repeated measures of slow-changing ecological features were made. Snapshots in time may be sufficient to capture the temporal dynamics of such phenomena (e.g. changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these the total time elapsed between the first and last measurements (the span of observation) may more closely approximate effective duration. However, many studies focused on more dynamic phenomena, and for these long periods of continuous observation may be more important for understanding dynamics than frequent repeats. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to directly measure fire behavior \cite{clements_observing_2007}, except patterns resolving themselves at scales greater than 12-24 hours. For this reason, and to provide a consistent standard for estimating duration, we recorded actual duration.

%Adding to our motivation to record actual observational scales was the fact that

\noindent \textbf{Broader implications}\\

Our results show that modern ecological observations most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).

\begin{enumerate}
\item General patterns of scales 
  \subitem{-} General patterns
  \subitem{-} The continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
  \subitem{-} Notable holes: high frequency (sub-daily, daily), high resolution

   \subitem{-} Duration hasn't improved much since Tilman (check this).
   \subitem{-} Evaluate how duration has improved since then. 387/749 of the papers he looked at from the journal Ecology (1977-1984) were mainly experimental or had an experimental component. 99 papers were mainly theoretical or conceptual. Remainder observational. We could compare our results to that here. Of his 623 field based studies (including 180 experiments), $\sim$41\% lasted less than one year,  $\sim$85\% less than three years, most studies $>$10 years and all $>$50 years were paleoecology/chronosequences.
  \subitem{-} Maybe mention here (or probably better later in the discussion), that our decision has the unfortunate consequence of not being directly comparable to previous assessments that included experiments. More of just a caveat or note for the reader who is familiar with the earlier work and their summary statistics
  \subitem But there do seem to be trends increasing extent (and duration, but need to put this above), perhaps due to increasing use of remote sensing 
\item Scales are not well-documented 
  \subitem Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}. (Scale is not a concern that cuts across the entire discipline of ecology).
